%------------------------------------------------------------------------
%                           DIFFEV Users guide
%------------------------------------------------------------------------

\documentclass[11pt]{report}

\usepackage[letterpaper]{geometry}
\geometry{height=8.5in}
\geometry{bottom=1.1in}
\geometry{left=1.1in}
\geometry{right=1.1in}

\usepackage[dvips]{graphicx}
\usepackage{fancyhdr}
\usepackage{floatflt}
\usepackage{tabularx}
\usepackage[round]{natbib}
\usepackage[hang,small,bf]{caption2}
\usepackage{fancyvrb}
\usepackage{scalefnt}
\usepackage{color}
\usepackage{url}
\usepackage{hyperref}
\usepackage[latin1]{inputenc}

\usepackage[T1]{fontenc}
\usepackage{palatino}
\input{../version.tex}
%\newcommand{\version}{5.16}

%------------------------------------------------------------------------

\newcommand{\Discus}{\textsc{Discus\ }}
\newcommand{\discus}{\textsc{Discus}}
\newcommand{\Diffev}{\textsc{Diffev\ }}
\newcommand{\diffev}{\textsc{Diffev}}
\newcommand{\Kuplot}{\textsc{Kuplot\ }}
\newcommand{\kuplot}{\textsc{Kuplot}}
\newcommand{\Suite}{\textsc{Discus Suite\ }}
\newcommand{\suite}{\textsc{Discus Suite}}

% General page style

\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\bfseries \slshape\leftmark}
\fancyhead[R]{\bfseries Users Guide}
\fancyfoot[R]{\bfseries Page \thepage}
\fancyfoot[L]{\bfseries \Diffev \version}
\renewcommand{\footrulewidth}{0.4pt}
\renewcommand{\headrulewidth}{0.4pt}

% Style for chapter start pages ..

\fancypagestyle{plain} {
  \fancyhf{}
  \fancyhead[R]{\bfseries Users Guide}
  \fancyfoot[L]{\bfseries \Diffev \version}
  \fancyfoot[R]{\bfseries Page \thepage}
  \renewcommand{\headrulewidth}{0.4pt}
}

% MacVerbatim style for macros

\definecolor{DarkBlue}{rgb}{0.1,0.0,0.5}

\DefineVerbatimEnvironment%
{MacVerbatim}{Verbatim}
{fontfamily=courier,
 fontsize=\scriptsize,
 fontseries=b,
 formatcom=\color{DarkBlue}}

\graphicspath{{pic/}}

\setlength\parindent{0pt}

%------------------------------------------------------------------------

\begin{document}

%------------------------------------------------------------------------
% Title
%------------------------------------------------------------------------

\begin{titlepage}
\begin{flushright}

  \hrule
  \vspace{15mm}
  \textbf{{\scalefont{8} \diffev}} \\
  \vspace{15mm}
  \textbf{{\scalefont{4} Users Guide}} \\
  \vspace{10mm}
  \textbf{{\Huge Version  \version}} \\
  \vspace{10mm}

  \hrule
  \vspace{45mm}
  \textbf{written by} \\

  \vspace{5mm}
  \textbf{\Large Reinhard Neder} \\
  Email: reinhard.neder@fau.de \\

  \vspace{20mm}
  \textbf{\Large \url{http://tproffen.github.io/DiffuseCode}} \\

  \vspace{3mm}
  \hrule

\end{flushright}
\begin{flushright}
  \textit{Document created: \today}
\end{flushright}

\end{titlepage}

%------------------------------------------------------------------------
% Preface
%------------------------------------------------------------------------

\chapter*{Preface}
\subsection*{Disclaimer}

The \Diffev software described in this guide is
provided without warranty of any kind.  No liability is taken for any loss
or damages, direct or indirect, that may result through the use of {\it
DIFFEV}.  No warranty is made with respect to this manual, or the program
and functions therein.  There are no warranties that the programs are free
of error, or that they are consistent with any standard, or that they will
meet the requirement for a particular application.  The programs and the
manual have been thoroughly checked.  Nevertheless, it can not be
guaranteed that the manual is correct and up to date in every detail. This
manual and the \Diffev program may be changed without notice.\par

\Diffev is intended as a public domain program. It may be used free
of charge.  Any commercial use is, however, not allowed without permission
of the authors.

%------------------------------------------------------------------------

\subsection*{Using DIFFEV}

Publications of results totally or partially obtained using the program
\Diffev should state that \Diffev was used and contain the
following reference:

\begin{quote}
  {\sc Neder, R.B.} in preparation - check website.
\end{quote}

%------------------------------------------------------------------------

\subsection*{More information}

This users guide can only provide program specific details. A
broader discussion of simulation techniques and some \Diffev
examples and macro files can be found in our book

\begin{quote}
  {\sc Neder, R.B. \& Proffen, Th.} "Diffuse Scattering and Defect Structure
  Simulations - A cook book using the programs DISCUS", {\it IUCr Texts on
  Crystallography}, Oxford University Press, 2007.
\end{quote}

%------------------------------------------------------------------------
% Table of contents, List of Tables etc..
%------------------------------------------------------------------------

\tableofcontents
%\listoffigures
%\listoftables

%------------------------------------------------------------------------
% Here all the chapter are included ...
%
% Alter or comment the \includeonly command above to work with the
% complete manual or only certain parts.
%
%------------------------------------------------------------------------

\include{dev_int}
\include{dev_evo}
\include{dev_exa}

%------------------------------------------------------------------------
% Appendices
%------------------------------------------------------------------------

\appendix
\include{appl_dif}

%------------------------------------------------------------------------
% Bibliography
%------------------------------------------------------------------------

\addcontentsline{toc}{chapter}{Bibliography}
\bibliography{diffev_man}
\bibliographystyle{kristal}

\end{document}
