%------------------------------------------------------------------------
%                           PDFFIT Users guide
%------------------------------------------------------------------------
% Version:  April 2007
%------------------------------------------------------------------------

\documentclass[11pt]{report}

\usepackage[letterpaper]{geometry}
\geometry{height=8.5in}
\geometry{bottom=1.1in}
\geometry{left=1.1in}
\geometry{right=1.1in}

\usepackage[dvips]{graphicx}
\usepackage{fancyhdr}
\usepackage{floatflt}
\usepackage{tabularx}
\usepackage[round]{natbib}
\usepackage[hang,small,bf]{caption2}
\usepackage{fancyvrb}
\usepackage{scalefnt}
\usepackage{color}
\usepackage{url}

\usepackage[T1]{fontenc}
\usepackage{palatino}

\newcommand{\version}{1.4}

% General page style

\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\bfseries \slshape\leftmark}
\fancyhead[R]{\bfseries Users Guide}
\fancyfoot[R]{\bfseries Page \thepage}
\fancyfoot[L]{\bfseries PDFFIT \version}
\renewcommand{\footrulewidth}{0.4pt}
\renewcommand{\headrulewidth}{0.4pt}

% Style for chapter start pages ..

\fancypagestyle{plain} {
  \fancyhf{}
  \fancyhead[R]{\bfseries Users Guide}
  \fancyfoot[L]{\bfseries PDFFIT \version}
  \fancyfoot[R]{\bfseries Page \thepage}
  \renewcommand{\headrulewidth}{0.4pt}
}

% MacVerbatim style for macros

\definecolor{DarkBlue}{rgb}{0.1,0.0,0.5}

\DefineVerbatimEnvironment%
{MacVerbatim}{Verbatim}
{fontfamily=courier,
 fontsize=\scriptsize,
 fontseries=b,
 formatcom=\color{DarkBlue}}

\graphicspath{{pic/}}

\setlength\parindent{0pt}

%------------------------------------------------------------------------

\begin{document}

%------------------------------------------------------------------------
% Title
%------------------------------------------------------------------------

\begin{titlepage}
\begin{flushright}

  \hrule
  \vspace{15mm}
  \textbf{{\scalefont{8} PDFFIT}} \\
  \vspace{15mm}
  \textbf{{\scalefont{4} Users Guide}} \\
  \vspace{10mm}
  \textbf{{\Huge Version  \version}} \\
  \vspace{10mm}

  \hrule
  \vspace{35mm}
  \textbf{written by} \\

  \vspace{5mm}
  \textbf{\Large Thomas Proffen} \\
  Email: tproffen@lanl.gov \\

  \vspace{ 2mm}
  \textbf{\Large Simon Billinge} \\
  Email: billinge@pa.msu.edu \\

  \vspace{12mm}
  \textbf{\Large \url{http://discus.sourceforge.net/}} \\

  \vspace{3mm}
  \hrule

\end{flushright}
\begin{flushright}
  \textit{Document created: \today \\ LA-UR 05-2800}
\end{flushright}

\end{titlepage}

%------------------------------------------------------------------------
% Preface
%------------------------------------------------------------------------

\chapter*{Preface}
\subsection*{Disclaimer}

The {\it PDFFIT} software described in this guide is
provided without warranty of any kind.  No liability is taken for any loss
or damages, direct or indirect, that may result through the use of {\it
PDFFIT}.  No warranty is made with respect to this manual, or the program
and functions therein.  There are no warranties that the programs are free
of error, or that they are consistent with any standard, or that they will
meet the requirement for a particular application.  The programs and the
manual have been thoroughly checked.  Nevertheless, it can not be
guaranteed that the manual is correct and up to date in every detail. This
manual and the {\it PDFFIT} program may be changed without notice.\par

{\it PDFFIT} is intended as a public domain program. It may be used free
of charge.  Any commercial use is, however, not allowed without permission
of the authors.

%------------------------------------------------------------------------

\subsection*{Acknowledgments}

The basic concept of the full profile refinement of the
pair distribution function (PDF) was adapted from the program {\it RESPAR}
written by Simon Billinge \citep{billi;b;lsfd98}. The routines for the
processing of the command language and the FORTRAN style interpreter are
taken from the program package {\it DISCUS} by Reinhard Neder and Thomas
Proffen \citep{prne97}. {\it PDFFIT} is distributed as part of that package
as well as a stand-alone program. The routines for the line editing were
taken from the program GNUPLOT.

%------------------------------------------------------------------------

\subsection*{Using PDFFIT}

Publications of results totally or partially obtained using the program
{\it PDFFIT} should state that {\it PDFFIT} was used and contain the
following reference:

\begin{quote}
  {\sc Proffen, Th. \& Billinge, S.J.L.} (1999) "PDFFIT, a Program
  for Full Profile Structural Refinement of the Atomic Pair
  Distribution Function" {\it J. Appl. Cryst.}, {\bf 32}, 572-575
\end{quote}

%------------------------------------------------------------------------
% Table of contents, List of Tables etc..
%------------------------------------------------------------------------

\tableofcontents

%------------------------------------------------------------------------
% Here all the chapter are included ...
%
% Alter or comment the \includeonly command above to work with the
% complete manual or only certain parts.
%
%------------------------------------------------------------------------

\include{pdf_int}
\include{pdf_qui}
\include{pdf_fil}
\include{pdf_cal}
\include{pdf_f77}
\include{pdf_fit}
\include{pdf_exa}
\include{pdf_mis}

%------------------------------------------------------------------------
% Appendices
%------------------------------------------------------------------------

\appendix
\include{pdf_apa}
\include{appl_pdf}

%------------------------------------------------------------------------
% Bibliography
%------------------------------------------------------------------------

\addcontentsline{toc}{chapter}{Bibliography}
\bibliography{pdf_man}
\bibliographystyle{kristal}

\end{document}
