%------------------------------------------------------------------------
%                           MIXSCAT Users guide
%------------------------------------------------------------------------

\documentclass[12pt]{report}

\usepackage[letterpaper]{geometry}
\geometry{height=8.5in}
\geometry{bottom=1.1in}
\geometry{left=1.1in}
\geometry{right=1.1in}

\usepackage[dvips]{graphicx}
\usepackage{fancyhdr}
\usepackage{floatflt}
\usepackage{tabularx}
\usepackage[round]{natbib}
\usepackage[hang,small,bf]{caption2}
\usepackage{fancyvrb}
\usepackage{scalefnt}
\usepackage{color}
\usepackage{url}
\usepackage{hyperref}

\usepackage[T1]{fontenc}
\usepackage{palatino}

\newcommand{\version}{1.0}
\newcommand{\mixscat}{\textsc {Mixscat}}
\newcommand{\Mixscat}{\textsc {Mixscat\ }}

% General page style

\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\bfseries \slshape\leftmark}
\fancyhead[R]{\bfseries Users Guide}
\fancyfoot[R]{\bfseries Page \thepage}
\fancyfoot[L]{\bfseries \Mixscat \version}
\renewcommand{\footrulewidth}{0.4pt}
\renewcommand{\headrulewidth}{0.4pt}

% Style for chapter start pages ..

\fancypagestyle{plain} {
  \fancyhf{}
  \fancyhead[R]{\bfseries Users Guide}
  \fancyfoot[L]{\bfseries \Mixscat \version}
  \fancyfoot[R]{\bfseries Page \thepage}
  \renewcommand{\headrulewidth}{0.4pt}
}

% MacVerbatim style for macros

\definecolor{DarkBlue}{rgb}{0.1,0.0,0.5}

\DefineVerbatimEnvironment%
{MacVerbatim}{Verbatim}
{fontfamily=courier,
 fontsize=\scriptsize,
 fontseries=b,
 formatcom=\color{DarkBlue}}

\graphicspath{{pic/}}

\setlength\parindent{0pt}

%------------------------------------------------------------------------

\begin{document}

%------------------------------------------------------------------------
% Title
%------------------------------------------------------------------------

\begin{titlepage}
\begin{flushright}

  \hrule
  \vspace{15mm}
  \textbf{{\scalefont{8} \mixscat}} \\
  \vspace{15mm}
  \textbf{{\scalefont{4} Users Guide}} \\
  \vspace{10mm}
  \textbf{{\Huge Version  \version}} \\
  \vspace{10mm}

  \hrule
  \vspace{20mm}
  \textbf{developed by} \\

  \vspace{5mm}
  \textbf{\Large Caroline Wurden} \\
  \vspace{2mm}
  \textbf{\Large Katharine Page} \\
  \vspace{2mm}
  \textbf{\Large Anna Llobet} \\

  \vspace{2mm}
  \textbf{\Large Thomas Proffen} \\
  \vspace{5mm}
  Email: tproffen@ornl.gov \\

  \vspace{12mm}
  \textbf{\Large \url{http://tproffen.github.io/DiffuseCode}} \\

  \vspace{3mm}
  \hrule

\end{flushright}
\begin{flushright}
  \textit{Document created: \today}
\end{flushright}
\end{titlepage}

%------------------------------------------------------------------------
% Preface
%------------------------------------------------------------------------

\chapter*{Preface}
\subsection*{Disclaimer}

The \Mixscat software described in this guide is
provided without warranty of any kind.  No liability is taken for any loss
or damages, direct or indirect, that may result through the use of \mixscat.  No warranty is made with respect to this manual, or the program
and functions therein.  There are no warranties that the programs are free
of error, or that they are consistent with any standard, or that they will
meet the requirement for a particular application.  The programs and the
manual have been thoroughly checked.  Nevertheless, it can not be
guaranteed that the manual is correct and up-to-date in every detail. This
manual and the \Mixscat program may be changed without notice.\par

\Mixscat is intended as a public domain program.  It may be used free
of charge.  Any commercial use is, however, not allowed without the
explicit written permission of the authors.

%------------------------------------------------------------------------

\subsection*{Acknowledgments}

The list of structure factors was adapted from the program
Lazy-Pulverix written by K.  Yvon, W.  Jeitschko \& E.  Parthe.

%------------------------------------------------------------------------

\subsection*{Using \mixscat}

Publication of results totally or partially obtained using the program \Mixscat should state that \Mixscat was used and contain the following
reference:

\begin{quote}
  {\sc Wurden, C., Page, K.L., Llobet, A. \& Proffen, Th.} (2010)
  "Extracting Differential Pair Distribution Functions using \mixscat"
  {\it J. Appl. Cryst.}, {\bf xx}, xxx--xxx
\end{quote}


%------------------------------------------------------------------------
% Table of contents, List of Tables etc..
%------------------------------------------------------------------------

\tableofcontents

%------------------------------------------------------------------------
% Here all the chapter are included ...
%------------------------------------------------------------------------

\include{mix_int}
\include{mix_exa}

%------------------------------------------------------------------------
% Appendices
%------------------------------------------------------------------------

\appendix
\include{mix_app}
\include{appl_mix}

%------------------------------------------------------------------------
% Bibliography
%------------------------------------------------------------------------

\addcontentsline{toc}{chapter}{Bibliography}
\bibliography{mix_man}
\bibliographystyle{kristal}

\end{document}
